---
title: "NBA Home-court Advantage"
author: "Renee Watson"
date: "2026-09-02"
output:
  html_document: default
  pdf_document: default
subtitle: 'DATA 607: Week 1 Assignment'
---

# NBA Home-court Advantage Analysis

## Approach

In this assignment , I selected a NBA dataset from FiveThirtyEight that contains information about NBA games, including the teams, game location, points scored, and game results. I plan to use this dataset to explore home-court advantage in the NBA.

My plan is to load data into R and select variables related to the team, opponent, game location, points scored, and whether the team won or lost. I will clean and rename variables as needed and compare games results from teams playing at home versus away.

Data Source: <https://raw.githubusercontent.com/reneewatson15/NBA-home-court-advantage-/refs/heads/main/nbaallelo.csv>

<https://data.fivethirtyeight.com>

## Codebase

```{r}
nba <- read.csv("https://raw.githubusercontent.com/reneewatson15/NBA-home-court-advantage-/refs/heads/main/nbaallelo.csv")

str(nba)
```


For this analysis, I selected variables related to the season, teams, game location, points scored, and game result. 

```{r}
nba_subset <- nba[c("year_id", "date_game", "team_id", "opp_id", "pts", "opp_pts", "game_location", "game_result")]

head(nba_subset)
```

I changed the names of the variables.
```{r}
names(nba_subset) <- c("season", "game_date", "team", "opponent", "team_points", "opponent_points", "game_location", "game_result")

head(nba_subset)
```


## Analysis

To examine home-court advantage, I selected games where the team was playing at home. 

```{r}
home_games <- nba_subset[nba_subset$game_location == "H", ]

head(home_games)
```

```{r}
home_results <- table(home_games$game_result)

home_results
```

```{r}
barplot(home_results,
        main = "NBA Home Game Results",
        xlab = "Game Result",
        ylab = "Number of Games")
```

In the plot, "L" represents a loss and "W" represents a win. Home teams won more games than they lost. 

```{r}
home_win_percentage <- mean(home_games$game_result == "W") * 100

round(home_win_percentage, 1)
```

The results show that NBA teams won 62.3% of games when playing at home. 

## Conclusion 

The results of this analysis show that NBA teams won 62.3% of their games when playing at home, suggesting that home-court advantage is present in the historical data. To extend this analysis, I could compare home winning percentages across different NBA seasons or individual teams. I could also use more recent data to determine whether home-court advantage has changed over time. 